\documentclass[]{note}
\usepackage{CJKutf8}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{braket}
\usepackage[ruled,vlined]{algorithm2e}
\def\redd#1{\textcolor{red}{#1}}
\def\dom{\text{dom}}
\def\ri{\mathbf{Relint}}
\title{{{On determinantal variety}}}

\author{{{Li Enyu}}}

\date{}

\begin{document}

\maketitle
\def\sym{\mathrm{Sym}}
\def\St{\mathrm{St}}
\def\H{\mathscr H}
\def\deter{\mathbb R_{\le r}^{m\times n}}
\section{The determinantal variety}The determinantal variety $\mathbb R_{\le r}^{m\times n}$ consists of matrices $\R^{m\times n}$ with rank less than or equal to $r$. The natural action of $O(m)\times O(n)$ gives an orbit decomposition \gs{
    \deter=\bigcup_{s\le r}\R_s^{m\times n}
} where each orbit is the smooth manifold consists of matrice of fixed rank (There is also an explicit cell decomposition but we do not need here).
Now  throughout this article let's fix a matrice with rank $s\le r$ in an SVD coordinate \gs{
    X=\jz{
        \Lambda&0\\0&0
    }
}where $\Lambda$ is diagonal of order $r$. To visualize, note that a neighborhood of $X$ lies in $\bigcup_{k\ge s}\R_k^{m\times n}$, namely there is a manifold part that preserving the rank, but also some directions that increasing rank, and the former forms the boundary of the latter. 
Consider the diffeomorphism near a neighborhood of $X$\gs{
\phi:U\to V\times \R_{\le r-s}^{(m-s)\times (n-s)}\subset\R^{s(m+n-s)}\times \R_{\le r-s}^{(m-s)\times (n-s)},\jz{A&B\\C&D}\mapsto \jz{A&B\\C& D/A}
}Here $D/A=D-CA^{-1}B$ is the Schur complement.
This implies locally $\R_{\le r}^{m\times n}$ \textbf{is a product of the manifold part and singular part}. Also note $d\phi$ is identity at $X$.

The tangent space $T_X\R^{m\times n}_s$ consists of matrices \gs{
    \jz{*&*\\ *&0}
}(
Throughout this article $*$ means arbitrary matrices.)
But the total tangent cone in $\deter$ consists of \gs{
    Y=\jz{*&*\\ *&D}, \text{rank}{D}\le r-s
}

\section{The invariant manifold}
Now consider a constraint $h:\R^{m\times n}\to \R^k$ that has constant rank and is invariant under the right action of $O(n)$. This implies the regular level set $\H=h^{-1}(0)$ has following properties:
\begin{enumerate}
    \item $\H$ is invariant under action of $O(n)$. 

    \item The stablizer $G_X$ acts on $N_X\H$ by identity. 
\end{enumerate}
These are \textbf{the only qualifications} we need. (In fact for transverality, 2 is enough)
The action of $O(n)$ gives a fundamental vector field on $\R_s^{m\times n}$, at $X$ this gives following tangent vector:
\gs{
    Y_1=\jz{
        \Lambda&0\\0&0
    }\Omega_0= \jz{
        \Lambda \Omega&*\\0&0
    }, \oo_0\in \mathfrak{so}_n,\oo\in \mathfrak{so}_s
}(i.e. here $\Omega_0$ and $\oo$ are skew-symmetric matrices.)

Also note the stablizer \gs{
    \sigma=\jz{0&0\\0&-I}\in O(n)_X
}
will flip the vectors of the form \gs{
    Y_2=\jz{0&0\\0&*}
}but by our property (2), $\sigma$ acts on normal space by identity, so $Y_2$ is automatically in $T_XH$. 
To summarize the tangent cone contains at least\gs{
\jz{\Lambda\oo&*\\0&*}
}


Another way to see this is to note that $h$ is just an $m$-variable function of vectors $(v_1,\cdots,v_m)$ in $\R^n$ that is invariant under orthogonal transformation, thus the value of $h$ are uniquely determined by their Gram matrix $\hk{\braket{v_i,v_j}}$. 
In orther words, consider the map\gs{
   \pi: \R^{m\times n}\to \text{Sym}^+; X\mapsto XX^t
}then there is a factorization $h=\bar h\circ \pi$. To see smoothness, note that although
$\pi$ is not a submersion globally,  it is a submersion when restricted to each fixed rank $s$ ( It's actually a fiber bundle with fiber $\St(s,n)$):\gs{
    \pi_s: \R^{mn}_s\to \mathrm{Sym}^+_s
} so $h$, when resricted to each fixed rank $s$, actually factors thru this map smoothly. Now it's not hard to see the vectors \gs{Y_1+Y_2=\jz{\Lambda\oo&*\\0&*}} are exactly kernel of $d\pi_s$. 

One can also use Schwartz's theorem in invariant theory to get this result:
every $G$-invariant map is a smooth map of its polynomial invariants\cite{sch}. In the $O(n)$ case, the First Fundamental Theorem (FFT) for the orthogonal group 
 states that all polynomial invariants are polynomials of indices of $XX^t$\cite{weyl}.
  Thus there is a smooth factorization $\bar h:\mathrm{Sym}^+\to \R^k$.

Finally we give an explicit expression of the tangent space. Consider the manifold \gs{R^{ms}\times 0=\hk{\jz{*&0\\ *&0}}} which is transverse to $\H$. Let $\H^s=\H\cap (R^{ms}\times 0)$, then we can write\gs{
    T\H=\hk{\jz{H_1&*\\ H_2&*},(H_1,H_2)\in T_X\H^s}
}


  \section{Transverse intersection}
Recall in section 1 we have a local diffeomorphism $\phi:\R^{m\times n} \subset V\times \R^{(m-s)(n-s)}$ at $X$ that sends $U$ to $V\times \R^{(m-s)(n-s)}_{\le r-s}$ with $d_X\phi$ identity. Now we can flatten $\H$ in this coordinate. This is standard in geometry. For completeness we give a proof here.

The projection $q:\H\cap \yk{V\times \R^{(m-s)(n-s)}}\to \R^{(m-s)(n-s)}$ is a submersion at $X$. Let $p$ be a projection of $V$ onto $\ker (d_X q)=T_X\H\cap \yk{V\times 0}$ and $q$ be its complement. Consider \gs{
    \psi:\H\cap \yk{V\times \R^{(m-s)(n-s)}}\to \ker d_Xq\times \R^{(m-s)(n-s)}, (x,y)\mapsto (p(x),y)
}This is a chart of $\H$ s.t. it is locally a product. Now we know that under this chart, 
\gs{
     \R_{\le r}^{m\times n}=& V\times \R^{(m-s)(n-s)}_{\le r-s}=\hk{\jz{*&*\\ *&D}:\text{rank}(D)\le r-s}\\ \H=&\H^s\times \R^{(m-s)(n-s)}=\hk{\jz{ A&* \\ B&*}:(A,B)\in \H^s}
}where $\H^s$ is some manifold in $\R^{ms}$. Now its clear  that their intersection is transverse as they are transverse in the first component and at the second component there is an inclusion $\R^{(m-s)(n-s)}_{\le r}\subset \R^{(m-s)(n-s)}$. As a consequence
\gs{
    T_X\yk{\R_{\le r}^{m\times n}\cap \H}=&T_X\yk{\R_{\le r}^{m\times n}}\cap T_X\H
}
This immediately gives the Theorem 1 in \cite{pap}

\section{Designularization}
The space $\deter$ is stratified, and each strata fibers over $\mathrm{Gr}_r(n)$, but not globally. We can view them as tuples $(v_1,\cdots,v_m)$, and at a sigularity some vector crashes into the span of the rest. To blowup up the singularity we need to add information to recover the direction of the crash: i.e. we record the $r$-plane that contains the current plane. With this intuition in mind, consider\gs{
    M=\hk{(X,V)\in \R^{m\times n}\times \mathrm{Gr}_r(n):row(X)\subset V }\to \mathrm{Gr}_r(n)
}this is just the total space of the $m$-th sum of the tautological bundles $\gamma^1_n$ \gs{
\hom(\R^m,\gamma^1_n)=(\gamma^1_n)^{\oplus m}
} which is smooth with the fiber $\hom(\R^m,V)=\R^{mr}$.

The desingularization is given by the projector onto the fiber direction \gs{
    &\Phi:M\to \R^{m\times n}_{\le r}\\&[(A,AG),V_G]\mapsto (A,AG)
}Thus $M$ indeed gives us a reparametrization of the space $\R_{\le r}^{mn}$.
 Note that \gs{\Phi^{-1}(X)=\hk{V\in \mathrm{Gr}_r(n):row(X)\subset V}=\mathrm{Gr}_{r-s}(n-s)}and $\Phi$ is diffeomorphism on smooth part $\R_r^{mn}$


We use Schubert cell to give local chart of $\mathrm{Gr}_r(n )  $  and $M$. Locally a vector space $V$ is 1-1 corresponds to an echelon form, 
again WLG we look at a fixed point\gs{
    V_0=\jz{I_r&0\\ 0& 0}\in \mathrm{Gr}_r(n)
}
a local chart is
\gs{
   \hk{ V_G=\jz{I_r&G\\ 0&0}}\subset \mathrm{Gr}_r(n)
}
and the fiber at $V_G$ consists of \gs{
F_{V_G}=A\jz{I_r&G\\ 0&0}=\jz{A&AG}
}
Thus a chart of the total space is just \gs{(A,G)\mapsto [(A,AG),V_G]}

We also give an explicit correspondence of Schubert coordinate and the projection coordinate used in \cite{pap}: for $(I_r,G)$, perform an $LQ$-decomposition\gs{
    (I_r,G)=L_GQ_G
}
where $Q_G\in \mathrm{St}_r(n)$ is $r\times n$ with rows orthorgonal and $L_G$ is $r\times r$ lower triangular. Both smoothly rely on $G$. The matrice $Q_G^tQ_G\in \mathrm{Sym}_r^+(n)$ is then a projector onto $V_G$.


Let's fix a point in $M$: \gs{p=(X,V_0),X=\jz{I_s&0\\ 0&0}} 
The tangent space along the fiber is\gs{
   T_pF_{V_0}=\hk{ \jz{A&0}}
} 
and the total tangent space is\gs{
    T_pM=\hk{
        \fk{ \jz{A_{sr}&G_{s(n-r)}\\A_{(m-s)r}&0}
        ,
        \jz{0_r&G\\0&0}
        }
    }
}
here the $G_{s(n-r)}$ is first $s$ rows of $G$.


Now consider \gs{
    M_h=\yk{\H\times \mathrm{Gr}_r(n)}\cap M=\hk{(X,V)\in \H\times \mathrm{Gr}_r(n):row(X)\subset V }\to \mathrm{Gr}_r(n)
}
Using our computation of tangent space of $\H$, we know this intersection is transverse to the fiber, i.e. \gs{
    T_{p}(\H)+T_{p}(F_{V_0})=T_pM
}So $M_h \to\mathrm{Gr}_r(n)$ is a submersion. It also follows that $M_h$ is a smooth embedded manifold. Since $M_h$ is closed submanifold of Euclidien space, it is also complete. Also, since for any compact $K$, $\Phi^{-1}(K)$ is closed in $K\times \mathrm{Gr}_r(n)$, $\Phi$ must be proper (this already implies $\Phi$ is a fiber bundle by Ehresmann's lemma, we will give an explicit trivialization later). 

We now perform a $LQ$-decomposition to $(I_r,G)$:\gs{
    \jz{I_r&G\\ 0&I_{n-r}}=\jz{L_G&0\\ *&L_G'}{Q_G}
}where $L_G$ is lower-triangular and $Q_G$ is orthorgonal, both smoothly rely on $G$. Since $h$ is $Q_G$ invariant, we have $h(A(I_r,G))=0\iff h(A(L_G,0))=0 $.
 A local chart of $M_h$ near $p$ is then given by\gs{
    \hk{ \fk{\jz{A&AG_1},\jz{0&G\\0&0}},h\yk{{A}L_G,0}=0,G=(G_1,G_2)^t}
 } substitute $(A,G)\to (AL_G,G)=(\bar A,G)$, we have\gs{
    \hk{ \fk{\jz{\bar AL_G^{-1}&\bar AL_G^{-1}G_1},\jz{0&G\\0&0}},h\yk{{\bar A},0}=0,G=(G_1,G_2)^t}
 }

The tangent space of $M_h$ at $p$ is \gs{
   T_p(M_h)=\hk{ \yk{\jz{H_1&*&G_{s(n-r)}\\H_2&*&0 },\jz{0&G\\0&0}} (H_1,H_2)\in T_X\H^s}
}
This is different from \cite{pap} because we use different charts on Grassmanian.

\section{Riemannian optimization}
We can pullback the problem as \gs{
    \min_{M_h}\bar f=f\Phi
}
where \gs{\Phi:M_h\subset \R^{m\times n}\times \mathrm{Gr}_r(n)\to \R_{\le r}^{m\times n}}
But now we need to figure out the behavious of $\Phi$.
The pushforward of $\Phi$ is exactly the first component of tangent space. Thus $\Phi$ will not be a submersion if $s\neq r$: the right-lower part of $d\Phi(T_pM_h)$ is $\R^{(m-s)(r-s)}\times 0$, so the right-lower part of $T(\R^{m\times b}_{\le r}\cap\H)$ i.e. the cone $\R^{(m-s)\times (n-s)}_{\le r-s}$, is not fully represented. This gives $1\TO 1$ of theorem 3 in \cite{pap}.

But we can save this by consider the second order behavior. Locally near $p$, a trivialization is
\gs{\jz{AL_G^{-1}&BL_G^{-1}&AL_G^{-1}G_1+BL_G^{-1}G_2\\CL_G^{-1}&DL_G^{-1}&CL_G^{-1}G_1+DL_G^{-1}G_2},h\yk{\jz{A&B&0\\C&D&0}}=0}In particular, note that $h$ is submersion onto $\R^{m(n-s)}$ so $B,D$ are free coordinates, we can take $B=0,G_1=0$ \gs{
    \jz{AL_G^{-1}&0&0\\CL_G^{-1}&DL_G^{-1}&DL_G^{-1}G_2},h\yk{\jz{A&0&0\\C&D&0}}=0
}Here we can use implicit function theorem so that $A,C$ relies on $D$ and some free parameter, namely $A(D,*),C(D,*)$.
Notice that\gs{L_G=I_r+O(|G|^2) } 
Now consider the matrice ${\epsilon}D$ and $\frac1\epsilon G$. The Taylor expansion along this direction is\gs{
    \jz{\partial^2_DA\cdot D^2&0&0\\\partial^2_DC\cdot D^2&0&0}\epsilon^2+\jz{\p_D A\cdot D &0&0\\ \p_D C\cdot D&D&0}\epsilon+\jz{0&0&0\\0&0&D G_2}+o(|G|^2+|D|^2)
}
The second order condition\gs{(f\circ\Phi(\gamma))''=f''(\Phi(\gamma))'+f'(\Phi(\gamma))''\ge 0} together with the first order condition\gs{
    f'\yk{\jz{H_1&*&*\\H_2&D &0}}=0 ,(H_1,H_2)\in T_X\H^s
} 
implies\gs{
    &f''\yk{\jz{A&0&0\\C&D &0}\epsilon}+f'\yk{
    \jz{\partial^2_DA\cdot D^2&0&0\\\partial^2_DC\cdot D^2&0&0}\epsilon^2}+f'\yk{\jz{0&0&0\\0&0&D G_2}
    }\\=& f''\yk{\jz{A&0&0\\C&D &0}\epsilon}+f'\yk{
    \jz{\partial^2_DA\cdot D^2&0&0\\\partial^2_DC\cdot D^2&0&0}\epsilon^2}+f'\yk{\jz{0&0&0\\0&D&D G_2}
    }\\\ge & 0
    }
for every $\epsilon$. Since epsilon is arbitary, 
\gs{f'\yk{\jz{0&0&0\\0&D&D G_2}
}\ge 0
}But this is true for $-D$, so
\gs{f'\yk{\jz{0&0&0\\0&D&D G_2}
}= 0
}
The matrices $D(I_{r-s},G_2)$ form a dense subset of $\R_{\le r-s}^{m-s\times n-s}$ and 
in particular, 
has the same span as the right-lower part of the tangent cone of $\R_{\le r}^{m\times n}$. 
As a consequence, 
This give $2\TO 1$ of theorem 3.

\section{Computation of Riemannian derivatives}
The metric on Grassmanian is endowed by the metric of $\mathrm{{Sym}}^+$ \gs{
Q^t_GQ_G
}

\begin{thebibliography}{9}

\bibitem{pap} Yang, Y., Gao, B. \& Yuan, Yx. A space-decoupling framework for optimization on bounded-rank matrices with orthogonally invariant constraints. Math. Program. (2026).

\bibitem{sch} https://mathoverflow.net/questions/472581/a-variant-of-schwarzs-theorem-on-generators-of-smooth-g-invariant-functions

\bibitem{weyl} https://mathoverflow.net/questions/364423/invariant-theory-for-the-orthogonal-group-and-clifford-algebras
\end{thebibliography}

\end{document}